Thermodynamic fragments in magic state theories

### 2020-12-04 magic_paper.tex
Added distillation bound in the main text

We would now like to study the constraints coming solely from majorization in the unital fragment. 
We consider the problem of purifying $\NN$ copies of a noisy strange state $\noisysN$ into a smaller number of $\tt$ copies of a less noisy strange state $\rho(\epsilon')\too{\tt}$, with $\epsilon' \le \epsilon$ and $\tt \leq \nn$. 
Since the state $\frac{1}{3} \I$ is free in this fragment, the interconversion problem is unaffected if we tensor in copies of it. 
This turns out to simplify the Lorenz curve analysis, and so we focus on the process,
\begin{equation}\label{eq:sudist}
	\noisys\too{2\nn} \longrightarrow \noisys(\tt, \epsilon', \nn) \coloneqq  \noisys(\epsilon')\too{2\tt} \otimes \left( \frac{1}{3}\id \right)\too{2(n-t)},
\end{equation}
where all copies are even.
Since the initial state can be written in the form $\noisys(2\nn, \epsilon, 0)$, it suffices to construct the Lorenz curve $\lc{\rho}{\id/3}(x)$ for the generic noisy strange state $\noisys(\nn, \epsilon ,\tt)$.
The exact expressions for the coordinates of this Lorenz curve can take $2^3 = 8$ different forms, depending on whether the noise level $\epsilon$ is lower/higher than $3/7$, the parity of the number of copies $\NN$ is even/odd and the location relative to the curve peak is on the left hand side (LHS -- including the maximum) / right hand side (RHS).
They are given in full detail in~\cref{app:lcsu_technical}.

Here we focus on the case of LHS points on Lorenz curves of even copies and noise levels $\epsilon \leq 3/7$.
The state $\rho(2\nn, \epsilon ,2\tt)$ has a Wigner distribution that takes on values $w_\ii$, with associated multiplicities $\mlt_\ii$, where
\begin{align}
w_\ii &= \left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^{2\ii}\left( -\frac{1}{3} + \frac{4}{9}\epsilon \right)^{2(\nn-\ii)} \\
m_\ii &= 8^{2\ii}\binom{2\nn}{2\ii}
\end{align}
From this, it is readily seen that $L_\sigma(x)$ has elbows on the LHS of its peak located at points $x=x_\ii$, where
\begin{equation}
x_\ii = \Phi_+ (\ii, \nn , \frac{8}{9}),
\end{equation}
with $\ii = 0,1, \dots , n$ and where $\Phi_+(i,n,p)$ is an even-power cumulant distribution function given by
\begin{equation}
\Phi_+(\ii, \nn, p) \coloneqq \sum_{\dummy=0}^{\ii} \binom{2\nn}{2\dummy} p ^{2\ii} (1-p)^{2(\nn-\ii)}.
\end{equation}
The value of the Lorenz curve at an elbow is given by
\begin{equation}
\lc{\rho}{\id/3}(x_{\ii}) \equiv L_{\ii} =  \left( \frac{5}{3} - \frac{8}{9}\epsilon\ \right)^{2\nn} \Phi_+\left(i;\nn,\frac{12-4\epsilon}{15-8\epsilon}\right),
\end{equation}
and we define $(x_{-1}, L_{-1}) = (0,0)$.

The coordinates of any point on the Lorenz curve LHS can be expressed as convex sums of the neighbouring elbows,
\begin{align}
    x_{\ii\jj\kk} &= \left( 1-p_{\ii\jj\kk}\right) x_{\ii-1} + p_{\ii\jj\kk} x_{\ii} \label{eq:lcsu_xcoord}\\
    L_{\ii\jj\kk} &= \left( 1-p_{\ii\jj\kk} \right) L_{\ii-1} + p_{\ii\jj\kk} L_{\ii}, \label{eq:lcsu_lcoord}
\end{align}
where we use a three-index parametrisation for the coordinates, with $\ii$ the index of the elbow that follows, $\jj = 1,\dots,\mlt_\ii$ and $\kk=1,\dots,9^{2(\nn-\tt)}$.
Therefore, for the initial state $\noisys\too{2\nn}$ we have
\begin{equation}
	p_{\ii\jj\kk} = \frac{\jj}{8^{2\ii}\binom{2\nn}{2\ii}},
\end{equation}
with $\ii=0,\dots,\nn,\ \jj(\ii) = 1,\dots,8^{2\ii}\binom{2\nn}{2\ii}$ and $\kk = 1$.
For the target state $\noisys(\epsilon')\too{2\tt} \otimes \left( \frac{1}{3}\id \right)\too{2(\nn-\tt)}$,
\begin{equation}
	p_{\ii'\jj'\kk'} = \frac{\kk' + (\jj'-1)9^{2(\nn-\tt)}}{9^{2(\nn-\tt)} 8^{2\ii'}\binom{2\tt}{2\ii'}}
\end{equation}
with $\ii'=0,\dots,\tt,\ \jj'(\ii') = 1,\dots,8^{2\ii'}\binom{\tt}{2\ii'}$ and $\kk' = 1,\dots,9^{2(\nn-\tt)}$.

>>>>

We would now like to study the constraints coming solely from majorization in the unital fragment. 
We consider the problem of purifying $\nn$ copies of a noisy strange state $\noisysn$ into a smaller number of copies $\tt$ of a less noisy strange state $\noisys(\epsilon')\too{\tt}$, with $\epsilon' \le \epsilon$ and $\tt \leq \nn$. 
Since the state $\frac{1}{3} \id$ is free, the interconversion problem is unaffected if we tensor in copies of it. 
This turns out to simplify the Lorenz curve analysis, and so we focus on the process,
\begin{equation}\label{eq:sudist}
	\noisys\too{\nn} \longrightarrow \noisys(\tt, \epsilon', \nn - \nn') \coloneqq  \noisys(\epsilon')\too{\tt} \otimes \left( \frac{1}{3}\id \right)\too{(\nn-\tt)},
\end{equation}
where all copies $\nn, \tt, \nn - \tt$ are even.
Since the initial state can be written in the form $\noisys(\nn, \epsilon, 0)$, it suffices to construct the Lorenz curve $\lc{\noisys}{\id/3}(x)$ for the generic noisy strange state $\noisys(\nn, \epsilon ,\tt)$.
This curve is defined at $9^\nn$ points between $0$ and $1$.
The exact expressions for the coordinates of these points can take $2^3 = 8$ different forms, depending on whether the noise level $\epsilon$ is low/high (relative to $\frac{3}{7}$), the parity of the number of copies $\nn$ is even/odd and the location relative to the curve peak is on the left hand side (LHS -- including the maximum) / right hand side (RHS).
The full details are provided in~\cref{app:lcsu_technical}.

Here we focus on the case of Lorenz curves of even copies $\nn, \tt$ and low noise levels ($\epsilon \leq 3/7$).
The part of the curve on the LHS of its peak is constructed by the positive components of the state Wigner distribution that take values $w_\ii$, with associated multiplicities $\mlt_\ii$,
\begin{align}
	w_\ii &= \left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^{2\ii}\left( -\frac{1}{3} + \frac{4}{9}\epsilon \right)^{\nn-2\ii}, \\
	m_\ii &= 8^{2\ii}\binom{\nn}{2\ii},
\end{align}
where $i=0,\dots,\frac{n}{2}$.
From this, it is readily seen that $\lc{\rho}{\id/3}(x)$ has elbows on the LHS of its peak located at points $x=x_\ii$, where
\begin{equation}
	x_\ii = \Phi_+ \left( 2\ii, \nn , \frac{8}{9} \right),
\end{equation}
with $\ii = 0,1, \dots , n$ and where $\Phi_+(i,n,p)$ is an even-power cumulant distribution function given by
\begin{equation}
	\Phi_+(2\ii, \nn, p) \coloneqq \sum_{\dummy=0}^{\ii} \binom{\nn}{2\dummy} p ^{2\dummy} (1-p)^{\nn-2\dummy}.
\end{equation}
The value of the Lorenz curve at an elbow is given by
\begin{equation}
	\lc{\rho}{\id/3}(x_{\ii}) \equiv L_{\ii} =  \left( \frac{5}{3} - \frac{8}{9}\epsilon\ \right)^{\nn} \Phi_+\left(2i;\nn,\frac{12-4\epsilon}{15-8\epsilon}\right),
\end{equation}
and we define $(x_{-1}, L_{-1}) = (0,0)$.

\iffalse
The coordinates of any point on the Lorenz curve LHS can be expressed as convex sums of the neighbouring elbows,
\begin{align}
    x_{\ii\jj\kk} &= \left( 1-p_{\ii\jj\kk}\right) x_{\ii-1} + p_{\ii\jj\kk} x_{\ii} \label{eq:lcsu_xcoord}\\
    L_{\ii\jj\kk} &= \left( 1-p_{\ii\jj\kk} \right) L_{\ii-1} + p_{\ii\jj\kk} L_{\ii}, \label{eq:lcsu_lcoord}
\end{align}
where we use a three-index parametrisation for the coordinates, with $\ii$ the index of the elbow that follows, $\jj = 1,\dots,\mlt_\ii$ and $\kk=1,\dots,9^{2(\nn-\tt)}$.
Therefore, for the initial state $\noisys\too{2\nn}$ we have
\begin{equation}
	p_{\ii\jj\kk} = \frac{\jj}{8^{2\ii}\binom{2\nn}{2\ii}},
\end{equation}
with $\ii=0,\dots,\nn,\ \jj(\ii) = 1,\dots,8^{2\ii}\binom{2\nn}{2\ii}$ and $\kk = 1$.
For the target state $\noisys(\epsilon')\too{2\tt} \otimes \left( \frac{1}{3}\id \right)\too{2(\nn-\tt)}$,
\begin{equation}
	p'_{\ii'\jj'\kk'} = \frac{\kk' + (\jj'-1)9^{2(\nn-\tt)}}{9^{2(\nn-\tt)} 8^{2\ii'}\binom{2\tt}{2\ii'}}
\end{equation}
with $\ii'=0,\dots,\tt,\ \jj'(\ii') = 1,\dots,8^{2\ii'}\binom{\tt}{2\ii'}$ and $\kk' = 1,\dots,9^{2(\nn-\tt)}$.
\fi

Lorenz curve comparison provides $9^\nn$ necessary constraints for the process~\cref{eq:sudist}.
\cref{thm:elbows} shows that they can be reduced to $\tt$ independent constraints.
However, in principle we can pick any location $x_\star$ at which the Lorenz curves are defined and we then get a necessary distillation constraint, 
\begin{equation}
	\lc{\noisys(\nn, \epsilon, 0)}{\frac{\id}{3}}(x_\star) \geq \lc{\noisys(\tt, \epsilon', \nn-\tt)}{\frac{\id}{3}}(x_\star).
\end{equation}

We choose $x_\star = x_0 = \Phi_+(0, \nn, \frac{8}{9}) = \frac{1}{9^{\nn}}$, the location of the first elbow of the initial state $\noisys(\nn, \epsilon, 0)$.
The initial state Lorenz curve coordinate at this location is
\begin{equation}
	L_\star = L_0 = \left( \frac{5}{3} - \frac{8}{9}\epsilon\ \right)^{\nn} \Phi_+\left(0;\nn,\frac{12-4\epsilon}{15-8\epsilon}\right) = \left(\frac{3-4\epsilon}{9}\right)^\nn
\end{equation}

The first elbow of the target state is located at $x'_0 = \frac{1}{9^{\tt}} < x_\star$, so in order to find the target state Lorenz curve coordinate $L'_\star$, we can always interpolate between the origin and the first target state elbow, so that
\begin{equation}
	L'_\star = \frac{x_\star}{x'_0}L_0 = \frac{1}{9^{\nn-\tt}}\Phi_+\left(0;\tt,\frac{12-4\epsilon}{15-8\epsilon}\right) = \frac{(3-4\epsilon)^{\tt}}{9^{\nn}}.
\end{equation}
The process in~\cref{eq:sudist} is therefore possible if $L_\star \geq L'_\star$ or equivalently,
\begin{align}
	\frac{\nn}{\tt} &\geq \frac{\ln{(3-4\epsilon')}}{\ln{(3-4\epsilon)}}, \\
	\epsilon &\leq \frac{3}{4} - \frac{1}{4}(3-4\epsilon')^{\frac{\tt}{\nn}}.
\end{align}
As an example, the process with $(\nn, \tt) = (12, 4)$ and $\epsilon' = 0.05$ yields $\epsilon \leq 0.398$.
Numerical comparison of the curves with varying $\epsilon$ suggests that we require $\epsilon \leq 0.373$ for the process to be possible.
\nick{All numerical examples I have tested, for various different cases, are quite close}
